---
title: 'Project 2: Modeling, Testing, and Predicting'
author: "SDS348"
date: ''
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

```{r setup, include=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

knitr::opts_chunk$set(echo = TRUE, eval = TRUE,fig.align="center",warning=FALSE,message=FALSE,fig.width=8, fig.height=5, linewidth=60)
options(tibble.width = 100,width = 100)
library(tidyverse)
```

# Modeling

## Instructions

A knitted R Markdown document (preferably HTML) and the raw R Markdown file (as .Rmd) should both be submitted to Canvas by 11:59pm on the due date. These two documents will be graded jointly, so they must be consistent (i.e., don’t change the R Markdown file without also updating the knitted document). In the .Rmd file for Project 2, you can copy the first code-chunk into your project .Rmd file to get better formatting. Notice that you can adjust the opts_chunk$set(...) above to set certain parameters if necessary to make the knitting cleaner (you can globally set the size of all plots, etc). You can copy the set-up chunk in Project2.Rmd: I have gone ahead and set a few for you (such as disabling warnings and package-loading messges when knitting)! 

Like before, I envision your written text forming something of a narrative structure around your code/output. All results presented must have corresponding code. Any answers/results/plots etc. given without the corresponding R code that generated the result will not be graded. Furthermore, all code contained in our project document should work properly. Please do not include any extraneous code or code which produces error messages. (Code which produces warnings is fine as long as you understand what the warnings mean.)

## Find data:

Find one dataset with at least 5 variables (ideally more!) that you want to use to build models/test hypotheses. At least one should be categorical (with 2-5 groups, ideally; definitely fewer than 10) and at least two should be numeric (taking on more than 10 distinct values). Ideally, at least of your variables will be binary (if not, you will have to create one by discretizing a numeric or collapsing levels of a categorical). You will need a minimum of 40 observations (*at least* 10 observations for every explanatory variable you have, ideally 20+ observations/variable).

It is perfectly fine to use either dataset (or the merged dataset, or a subset of your variables) from Project 1. However, I might encourage you to diversify things a bit and choose a different dataset to work with (particularly if the variables did not reveal interesting associations in Project 1 that you want to follow up with). The only requirement/restriction is that you may not use data from any examples we have done in class or lab. It would be a good idea to pick more cohesive data this time around (i.e., variables that you actually thing might have a relationship you would want to test). Think more along the lines of your Biostats project.

Again, you can use data from anywhere you want (see bottom for resources)! If you want a quick way to see whether a built-in (R) dataset has binary and/or character (i.e., categorical) variables, check out this list: https://vincentarelbundock.github.io/Rdatasets/datasets.html.


## Guidelines and Rubric

- **0. (5 pts)** Introduce your dataset and each of your variables (or just your main variables if you have lots) in a paragraph. What are they measuring? How many observations?

This dataset explores the energy produced by each US state in the year 2009. The energy produced is described by biomass, geothermal, solar, wind, hydro, HPS, coal, petroleum, gas, nuclear, and other. The states are separated into regions. This dataset was extremely robust, but I was interested in seeing the relationship between region of the US and the type of energy produced. The regions for this dataset are Alaska, California, Florida, Hawaii, Heartland, Mid Atlantic, Midwest, New England, New York, Southeast, Texas, and West. During the climate crisis, it is important to look to our shortcomings and our strengths. I aim to have a career in sustainable energy law, so it appealed to me to find a relationship between different types of energy production and their location in the US.

- **1. (15 pts)** Perform a MANOVA testing whether any of your numeric variables (or a subset of them, if including them all is unreasonable or doesn't make sense) show a mean difference across levels of one of your categorical variables (3). If they do, perform univariate ANOVAs to find response(s) showing a mean difference across groups (3), and perform post-hoc t tests to find which groups differ (3). Discuss the number of tests you have performed, calculate the probability of at least one type I error (if unadjusted), and adjust the significance level accordingly (bonferroni correction) before discussing significant differences (3). Briefly discuss some of the MANOVA assumptions and whether or not they are likely to have been met here (no need for anything too in-depth) (2).

```{r}


library(readxl)
Project2RawData <- read_xlsx("Project2RawData.xlsx")

glimpse(data)
library(mvtnorm)
library(ggExtra)

glimpse(Project2RawData)
library(mvtnorm); library(ggExtra)
df<-rmvnorm(1000,mean=c(0,0),sigma=matrix(c(1,.5,.5,1),ncol=2,byrow=T))
df<-data.frame(df)%>%rename(Y1=X1,Y2=X2)
p<-ggplot(df, aes(Y1,Y2))+geom_point(alpha=.5)+geom_density_2d(h=2)+coord_fixed()
ggMarginal(p,type="density",xparams = list(bw=.5), yparams=list(bw=.5))
p

ggplot(Project2RawData,aes(x=Petroleum,y=Gas))+geom_point(alpha=.5)+geom_density_2d(h=2)+coord_fixed()+facet_wrap(~Region)

library(rstatix)
group<-Project2RawData$Region
DVs<-Project2RawData%>%select(Biomass,Geothermal,Solar,Wind,Hydro,HPS,Coal,Petroleum,Gas,Nuclear,Other)
man1<-manova(cbind(Biomass,Geothermal,Solar,Wind,Hydro,HPS,Coal,Petroleum,Gas,Nuclear,Other)~Region,data=Project2RawData)
summary(man1)
summary.aov(man1)

Project2RawData%>%group_by(Region)%>%summarize(mean(Biomass),mean(Geothermal),mean(Solar),mean(Wind),mean(Hydro),mean(HPS),mean(Coal),mean(Petroleum),mean(Gas),mean(Nuclear),mean(Other))

```

The null hypothesis for a MANOVA test states that for each numeric response variable, the mean of all groups (regions of the US) are equal. The alternate hypothesis states that for at least 1 measure of energy produced, at least 1 region mean differs. This test is appropriate when we have multiple response variables (ie multiple types of energy produced in a year) to see if any differ by the categorical explanatory variable of US region. The assumptions for MANOVA are extensive. Random sampling, independent observations, multivariate normality of DVs, homogeneity of within group covariance matrices, linear relationships among DVs, no extreme univariate or multivariate outliers, and no mukticollinearity are required. This is a lot of assumptions, and it's unlikely that all of them are met in this case. Again, the regions for this dataset are Alaska, California, Florida, Hawaii, Heartland, Mid Atlantic, Midwest, New England, New York, Southeast, Texas, and West. I decided to focus on 6 numeric variables: wind, hydro, petroleum, solar, coal, and gas.

The overeall manova test was not significant. I was shocked by this! If everything had been significant, I would have had 11 response variables for 12 groups. That would mean 1 MANOVA, 11 ANOVA, and 12 t-tests, or 24 tests total. I included the code for the first steps to be safe.



- **2. (10 pts)** Perform some kind of randomization test on your data (that makes sense). The statistic can be anything you want (mean difference, correlation, F-statistic/ANOVA, chi-squared), etc. State null and alternative hypotheses, perform the test, and interpret the results (7). Create a plot visualizing the null distribution and the test statistic (3).

```{r}
Project2RawData<-na.omit(Project2RawData)
Project2RawData$Biomass%>%sort
mean(Project2RawData$Biomass)
samp1<-sample(Project2RawData$Biomass,replace=T)
sort(samp1)
samp1
mean(samp1)
means<-vector()
for(i in 1:5000){
samp<-sample(Project2RawData$Biomass,replace=T)
means[i]<-mean(samp)
}
quantile(means,c(0.025,0.975),na.rm=TRUE)
ggplot()+geom_histogram(aes(means))+geom_vline(xintercept=quantile(means,c(.025,.975)))
sd(means)
ggplot()+geom_histogram(aes(means))+
  geom_vline(xintercept=mean(means)+c(-1,1)*sd(means))



library(vegan)
dists<-Project2RawData%>%select(Biomass, Wind)%>%dist()
adonis(dists~Region,data=Project2RawData)

```
The null hypothesis is that the true mean of biomass generated in the US by state does not differ from the mean of the biomass generated in this dataset from 2009. The alternate hypothesis is that the true mean of biomass does differ significantly from the mean in this dataset. The histogram above shows a normal distribution with the quantiles included. Our mean falls within the range of sd. We can conclude that it is representative of the true mean of biomass, so we fail to reject the null.

- **3. (40 pts)** Build a linear regression model predicting one of your response variables from at least 2 other variables, including their interaction. Mean-center any numeric variables involved in the interaction.

```{r}

Project2RawData%>%ggplot(aes(Nuclear,Coal))+geom_point()
fit<-lm(Nuclear~Coal+Gas,data=Project2RawData);summary(fit)

resCoal<-lm(Nuclear~Gas,data=Project2RawData)$residuals
resGas<-lm(Nuclear~Coal,data=Project2RawData)$residuals
coef(lm(resCoal~resGas))
coef(lm(Nuclear~Coal+Gas,data=Project2RawData))

summary(fit)
Nuclear<-Project2RawData$Nuclear
(sum((Nuclear-mean(Nuclear))^2)-sum(fit$residuals^2))/sum(Nuclear-mean(Nuclear)^2)

library(sandwich);library(lmtest)
fit<-lm(Nuclear~Coal+Gas,data=Project2RawData)
bptest(fit)
summary(fit)$coef[,1:2]
coeftest(fit,vcov=vcovHC(fit))[,1:2]


fit<-lm(Nuclear~Coal*Gas,data=Project2RawData)
summary(fit)
Project2RawData$Coal_c<-Project2RawData$Coal-mean(Project2RawData$Coal)
Project2RawData$Gas_c<-Project2RawData$Gas-mean(Project2RawData$Gas)
fit<-lm(Nuclear~Coal_c*Gas_c,data=Project2RawData)
summary(fit)
resids<-fit$residuals;fitvals<-fit$fitted.values
ggplot()+geom_point(aes(fitvals,resids))+geom_hline(yintercept=0, col="red")
bptest(fit)
ggplot()+geom_histogram(aes(resids),bins=20)






```
I first ran the model in a simple linear model without interactions to get a general idea of how the data looked. 5.069x10^6 is the predicted value of nuclear energy production when coal and gas are 0. 2.43x10^-1 is the slope for coal on nuclear when holding gas constant. 1.26x10^-1 is the slope of gas on nuclear when holding coal constant. This linear model returned a massive residual standard error of 17960000 on 48 df. I could tell from an initial ggplot that heteroskedasticity was violated. I wanted to check, so I went ahead and did a Breusch-Pagan test, and adjusted.

I reran the model to account for interactions and mean center, as well as violations of assumptions. With interactions, 1.86x10^6 is the predicted value of nuclear energy production when coal and gas are 0. 0.308 is the slope for coal on nuclear when holding gas constant. 0.3012 is the slope of gas on nuclear when holding coal constant. After plotting the residuals, it was clear that homoskedasticity was violated. There were also many outliers that made it clear normality was violated as well. However, the bp test returned a p-value of 0.001187. 


    - Interpret the coefficient estimates (do not discuss significance) (10)
    - Plot the regression using `ggplot()` using geom_smooth(method="lm"). If your interaction is numeric by numeric, refer to code in the slides to make the plot or check out the `interactions` package, which makes this easier. If you have 3 or more predictors, just chose two of them to plot for convenience. (10)
    - What proportion of the variation in the outcome does your model explain? (4)
    - Check assumptions of linearity, normality, and homoskedasticity either graphically or using a hypothesis test (5)
    - Regardless, recompute regression results with robust standard errors via `coeftest(..., vcov=vcovHC(...))`. Discuss significance of results, including any changes from before/after robust SEs if applicable. (10)

- **4. (5 pts)** Rerun same regression model (with the interaction), but this time compute bootstrapped standard errors (either by resampling observations or residuals). Discuss any changes you observe in SEs and p-values using these SEs compared to the original SEs and the robust SEs)

```{r}
samp_distn<-replicate(5000, {
  boot_dat<-boot_dat<-Project2RawData[sample(nrow(Project2RawData),replace=TRUE),]
  fit<-lm(Nuclear ~ Coal_c * Gas_c, data=boot_dat)
  coef(fit)
})
samp_distn%>%t%>%as.data.frame%>%summarize_all(sd)
samp_distn %>% t %>% as.data.frame %>% pivot_longer(1:3) %>% group_by(name) %>%
summarize(lower=quantile(value,.025), upper=quantile(value,.975))
fit<-lm(Nuclear~Coal_c*Gas_c,data=Project2RawData)
resids<-fit$residuals
fitted<-fit$fitted.values
resid_resamp<-replicate(5000,{
    new_resids<-sample(resids,replace=TRUE)
    Project2RawData$new_y<-fitted+new_resids 
    fit<-lm(Nuclear~Coal_c*Gas_c,data=Project2RawData)
    coef(fit)
})

resid_resamp%>%t%>%as.data.frame%>%summarize_all(sd)
resid_resamp%>%t%>%as.data.frame%>%pivot_longer(1:3)%>%group_by(name)%>%
 summarize(lower=quantile(value,.025), upper=quantile(value,.975))



```
The bootstrapped values returned a big increase in the intercept and coal and gas values compared to the robust values. 


- **5. (30 pts)** Fit a logistic regression model predicting a binary variable (if you don't have one, make/get one) from at least two explanatory variables (interaction not necessary). 

    - Interpret coefficient estimates in context (10)
    - Report a confusion matrix for your logistic regression (5)
    - Compute and discuss the Accuracy, Sensitivity (TPR), Specificity (TNR), Precision (PPV), and AUC of your model (5)
    - Using ggplot, make a density plot of the log-odds (logit) colored/grouped by your binary outcome variable (5)
    - Generate an ROC curve (plot) and calculate AUC (either manually or with a package); interpret (5)

```{r}

data<-Project2RawData%>%mutate(y=ifelse(Region=="Texas",1,0))
head(data)
odds<-function(p)p/(1-p)
p<-seq(0,1,by=.1)
cbind(p, odds=odds(p))%>%round(4)
head(data)


fit<-glm(y~Petroleum+Gas,data=data,family=binomial(link="logit"))
fit
coeftest(fit)
exp(coef(fit))
exp(-4.96E15)

logistic<-function(x){exp(x)/(1+exp(x))}
ggplot()+geom_function(fun=logistic)+xlab("t")+ylab("p(y=1)")+xlim(-10,10)


data<-data%>%mutate(petro_cat=ntile(Petroleum,3))
data$petro_cat<-factor(data$petro_cat,labels=c("S","M","L"))
fit1<-glm(y~petro_cat,data=data,family=binomial(link="logit"))
summary(fit1)
exp(-18.793)
exp(-18.793)

info<-table(truth=data$y,prediction=data$petro_cat)%>%addmargins
info
(16)/51
(17)/50
(16/17)

exp(coef(fit1))
coef(fit1)%>%round(5)%>%data.frame
exp(-2.773)
odds2prob<-function(odds){odds/(1+odds)}
odds2prob(0.06247)
exp(-1.879)


probs<-predict(fit1,type="response")
table(predict=as.numeric(probs>0.5),truth=data$y)%>%addmargins
hist(table(predict=as.numeric(probs>0.5),truth=data$y)%>%addmargins
)
model.df=augment(fit1)%>%rename(log_odds='.fitted',)
glimpse(model.df)
ggplot(model.df,aes(y,log_odds,colour=y))+geom_line()+geom_smooth(se=TRUE)
model.df
ggplot(model.df, aes(x=log_odds , y=model.df$y )) +
  geom_point()  +  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial)) +  facet_wrap( ~ petro_cat)

mean(data[data$y==1,]$prob>.1)
mean(data[data$y==1,]$prob<.1)
 
logit<-glm(y~petro_cat,data=data,family=binomial(link="logit"))
logit
exp(coef(fit))
exp(cbind(coef(fit),confint(fit)))



```
Petroleum energy production increases the log odds of the region being Texas. For petroleum = 0, the log odds is -4.96x10^15, odds of region being Texas is e^-4.96x10^15, or ~0. For petroleum production = 1, the log odds is -4.96x10^15 + 5.237x10^8, or still roughly 0. Every one unit increase in petroleum energy production (note that 1 unit is incredibly small when dealing with this kind of variable) causes odds to multiply by 1424379. I reran this model dividing petroleum into 3 categorial levels. The model was  more successful this way. Odds of region being Texas for M are 6.89x10^9 times that of S. Odds of region being Texas for M are 6.89x10^9 times that of M.

Accuracy of the second model was 0.314. Specificity was 1.0. Sensitivity was 0.34. Precision was 0.941. The logit score was 0.1527. 

- **6. (25 pts)** Perform a logistic regression predicting the same binary response variable from *ALL* of the rest of your variables (the more, the better!) 

    - Fit model, compute in-sample classification diagnostics (Accuracy, Sensitivity, Specificity, Precision, AUC), and interpret (5)
    - Perform 10-fold (or repeated random sub-sampling) CV with the same model and report average out-of-sample classification diagnostics (Accuracy, Sensitivity, Specificity, Precision, and AUC); interpret AUC and compare with the in-sample metrics (10)
    - Perform LASSO on the same model/variables. Choose lambda to give the simplest model whose accuracy is near that of the best (i.e., `lambda.1se`). Discuss which variables are retained. (5)
    - Perform 10-fold CV using only the variables lasso selected: compare model's out-of-sample AUC to that of your logistic regressions above (5)

```{r}
fit3<-glm(y~Biomass+Geothermal+Solar+Wind+Hydro+HPS+Coal+Petroleum+Gas+Other,data=data,family="binomial")
coeftest(fit3)
prob3<-predict(fit3,type="response")
pred<-ifelse(prob3>.5,1,0)

exp(coef(fit3))%>%round(3)%>%t

class_diag <- function(probs,truth){
 tab<-table(factor(probs>.5,levels=c("FALSE","TRUE")),truth)
  acc=sum(diag(tab))/sum(tab)
  sens=tab[2,2]/colSums(tab)[2]
  spec=tab[1,1]/colSums(tab)[1]
  ppv=tab[2,2]/rowSums(tab)[2]
  f1=2*(sens*ppv)/(sens+ppv)
   if(is.numeric(truth)==FALSE & is.logical(truth)==FALSE) truth<-as.numeric(truth)-1
    ord<-order(probs, decreasing=TRUE)
  probs <- probs[ord]; truth <- truth[ord]
  TPR=cumsum(truth)/max(1,sum(truth))
  FPR=cumsum(!truth)/max(1,sum(!truth))
  dup<-c(probs[-1]>=probs[-length(probs)], FALSE)
  TPR<-c(0,TPR[!dup],1); FPR<-c(0,FPR[!dup],1)
  n <- length(TPR)
  auc<- sum( ((TPR[-1]+TPR[-n])/2) * (FPR[-1]-FPR[-n]) )
  data.frame(acc,sens,spec,ppv,f1,auc)
}
finaldat<-data%>%select(-State,-Abbreviation,-Region,-petro_cat)
head(finaldat)
fit<-glm(y~(.)^2,data=finaldat,family="binomial")
prob<-predict(fit,type="response")
class_diag(prob,finaldat$y)

set.seed(1234)
first_half<-finaldat%>%sample_frac(0.5)
second_half<-anti_join(finaldat,first_half)
fit<-glm(y~(.)^2,data=first_half,family="binomial")
probs1<-predict(fit,newdata=second_half,type="response")
truth1<-second_half$y
table(prediction=as.numeric(probs1>.5),truth1)

fit<-glm(y~(.)^2,data=second_half,family="binomial")
probs2<-predict(fit,newdata=first_half,type="response")
truth2<-first_half$y
table(prediction=as.numeric(probs2>.5),truth2)
class_diag(probs2,truth2)



set.seed(1234)
k=10
donedat<-finaldat%>%sample_frac
donedat
folds<-ntile(1:nrow(finaldat),n=10)
probl<-predict(fit,type="response")
class_diag(prob,donedat$y)

lastdat<-donedat%>%sample_frac
folds<-ntile(1:nrow(data),n=10)
diags<-NULL

library(glmnet)
y<-as.matrix(data$y)
y
x<-model.matrix(Region~.,data=data)[,-1]
head(x)
x<-scale(x)




```
The auc was quite high at 0.98, meaning that Texas region is being predicted fairly accurately. Acc and spec were quite high as well, while sense and ppv were low. 



## Where do I find data again?

You can choose ANY datasets you want that meet the above criteria for variables and observations. You can make it as serious as you want, or not, but keep in mind that you will be incorporating this project into a portfolio webpage for your final in this course, so choose something that really reflects who you are, or something that you feel will advance you in the direction you hope to move career-wise, or something that you think is really neat, or whatever. On the flip side, regardless of what you pick, you will be performing all the same tasks, so it doesn't end up being that big of a deal.

If you are totally clueless and have no direction at all, log into the server and type 

```{R eval=F}
data(package = .packages(all.available = TRUE))
```

This will print out a list of **ALL datasets in ALL packages** installed on the server (a ton)! Scroll until your eyes bleed! Actually, do not scroll that much... To start with something more manageable, just run the command on your own computer, or just run `data()` to bring up the datasets in your current environment. To read more about a dataset, do `?packagename::datasetname`. 

If it is easier for you, and in case you don't have many packages installed, a list of R datasets from a few common packages (also downloadable in CSV format) is given at the following website: https://vincentarelbundock.github.io/Rdatasets/datasets.html.

- A good package to download for fun/relevant data is `fivethiryeight`. 
Run `install.packages("fivethirtyeight"),` load the packages with `library(fivethirtyeight)`, run `data()`, and then scroll down to view the datasets. Here is an online list of all 127 datasets (with links to the 538 articles). Lots of sports, politics, current events, etc.

- If you have already started to specialize (e.g., ecology, epidemiology) you might look at discipline-specific R packages (vegan, epi, respectively). We will be using some tools from these packages later in the course, but they come with lots of data too, which you can explore according to the directions above

- However, you *emphatically DO NOT* have to use datasets available via R packages! In fact, I would much prefer it if you found the data from completely separate sources and brought them together (a much more realistic experience in the real world)! You can even reuse data from your SDS328M project, provided it shares a variable in common with other data which allows you to  merge the two together (e.g., if you still had the timestamp, you could look up the weather that day: https://www.wunderground.com/history/). If you work in a research lab or have access to old data, you could potentially merge it with new data from your lab!

- Here is a curated list of interesting datasets (read-only spreadsheet format): https://docs.google.com/spreadsheets/d/1wZhPLMCHKJvwOkP4juclhjFgqIY8fQFMemwKL2c64vk/edit

- Here is another great compilation of datasets: https://github.com/rfordatascience/tidytuesday

- Here is the UCI Machine Learning Repository: https://archive.ics.uci.edu/ml/index.php

    - See also https://en.wikipedia.org/wiki/List_of_datasets_for_machine-learning_research#Biological_data

- Here is another good general place to look: https://www.kaggle.com/datasets

- To help narrow your search down or to see interesting variable ideas, check out https://www.tylervigen.com/spurious-correlations. This is the spurious correlations website, and it is fun, but if you look at the bottom of each plot you will see sources for the data. This is a good place to find very general data (or at least get a sense of where you can scrape data together from)!

- If you are interested in medical data, check out www.countyhealthrankings.org

- If you are interested in scraping UT data, they make *loads* of data public (e.g., beyond just professor CVs and syllabi). Check out all the data that is available in the statistical handbooks: https://reports.utexas.edu/statistical-handbook

##### Broader data sources:

[Data.gov](www.data.gov) 186,000+ datasets!

[Social Explorer](Social Explorer) is a nice interface to Census and American Community Survey data (more user-friendly than the government sites). May need to sign up for a free trial.

[U.S. Bureau of Labor Statistics](www.bls.gov)

[U.S. Census Bureau](www.census.gov)

[Gapminder](www.gapminder.org/data), data about the world.

...





